Evolution of Flare Activity in GKM Stars Younger than 300 Myr over Five Years of TESS Observations

### 2024-01-22 src/tex/ms.tex
ready for submission!

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}
\usepackage{rsfso}

@@

% Abstract with filler text
\begin{abstract}
Stellar flares are short-duration ($<$\,hours) bursts of radiation associated with surface magnetic reconnection events. Stellar magnetic activity generally decreases as a function of both age and Rossby number, $R_0$, a measure of the relative dominance of the convective and rotational dynamos. However, without young stars have been overlooked in flare studies due to lack of long baseline observations and challenges with flare-detection methods. With this motivation, we select a sample of stars that are either (i) members, (ii) candidate members or (iii) have a high-probability of being  members of 26 nearby young moving groups, clusters, or associations all known to be less than 300~Myr old that have been observed by the Transiting Exoplanet Survey Satellite (TESS) at 2-minute cadence.  With this sample of \nstars\ stars younger than 300~Myr, we identified \nflares\ flares originating from \nflarestars\ stars. Moreover, we robustly measured the rotation periods of \nprot~stars. We measure the flare frequency distribution (FFD) slope as a function of both age and spectral type and find the FFD slope, $\alpha$, saturates for all spectral types at $\alpha = -0.5$ and is constant over the 300~Myr window studied here. Additionally, we find that $R_0$ and flare rate for stars $t_\textrm{age} = 100 - 300$~Myr are saturated out to $R_0 = 0.14$, which is consistent with other indicators of magnetic activity. We cross match our targets with \textit{GALEX} and find no correlation between flare rate and Far- and Near-Ultraviolet flux. We compare the flare rates of planet hosting stars to comparable, larger samples of stars and find the majority of planet-hosting stars are relatively flare quiet. This may impact the atmospheric evolution of young short-period exoplanets. Finally, we detect evidence of stellar flare variations across five years of TESS data in eleven young stars, likely as a result of longer timescale stellar cycles.

@@

Stellar flares are the radiation component of magnetic reconnection events \citep{benz10}. Such events are readily seen on the Sun \citep{carrington1859, lu91, fletcher11}, particularly now as we enter a maximum in the solar cycle \citep{webb94}. Such events can be used to understand the magnetic activity of other stars \citep{feigelson99, berdyugina05, kowalski10, feinstein22_criticality}. Additionally, by studying stellar flares we can understand how these short-duration events may impact exoplanet evolution \citep{france16, guenther19_flares, chen21}. While we cannot angularly resolve stellar flare events on other stars, we can detect and characterize both spectroscopic and photometric signatures of such events. Spectroscopic characterization of stellar flares inform our understanding of non-thermal processes affiliated with such events, e.g. coronal mass ejections \citep{argiroffi19, vida19} or proton beams \citep{orrall76, woodgate92}.

Photometric observations of stars are more readily available thanks to exoplanet transit discovery missions, and allow us to statistically characterize flare rates and energies. NASA's \textit{Kepler}, \textit{K2}, and the Transiting Exoplanet Survey Satellite (TESS) missions have provided a wealth of stellar variability data in addition to their primary objective of detecting transiting exoplanets.  Within time-series photometry, stellar flares can be identified by a sharp rise in the stellar flux and a subsequent exponential decay, corresponding to the cooling rate \citep{}. \textit{Kepler} \citep{Borucki10} provided long-baseline high-cadence observations used to identify stellar flares. There has been extensive studies of flares in \textit{Kepler} data, from the statistics of superflares on solar-type stars \citep[e.g.][]{notsu13, shibayama13, maehara15, okamoto21} to low-mass stars \citep[e.g.][]{Hawley14, silverberg16}. \cite{davenport19} found that flare activity decreased with increasing rotation period for 347 GKM stars. However, the flare frequency distribution (FFD) slope did not change significantly as a function of age. As a caveat, the ages of the stars were determined based on their rotation periods, relying on the assumption that gyrochronology alone accurately ages stars.

\textit{K2} provided 70-day baseline observations for a handful of young stars in groups such as Upper Scorpius,  Pleiades, Hyades, and Praespe clusters. \cite{ilin19, ilin21} analyzed flares K and M stars in these clusters and found that the overall flare activity decreased as a function of age. This relationship was steeper for more massive stars in the sample. \cite{paudel18} surveyed 10 M6 - L0 dwarfs observed with \textit{K2} and found the L0 dwarf had a much shallower FFD than the M dwarfs. On average, they found that very young targets, defined by the tangential velocity of the star, overall had more flares.

More recently, TESS \citep{ricker15} has provided near all-sky photometric observations at 30-minute cadence or less. Such an observing strategy has allowed for more detailed studies of young stellar flares from nearby young moving groups and associations, which are disperse in the sky \citep{gagne18}. These new data have allowed for detailed studies of individual stars, for example characterizing eight superflares ($E_f > 10^{34}$~erg) on the young star AB Doradus over $\sim 60$ days of continuous observations \citep{schmitt19}.

The 11-year solar activity cycle represents a change in the magnetic activity of the Sun, and manifests itself in a variety of observables e.g., an increase in the total number of sunspots \citep{clette14, kilcik14}, an increase in the number of flares and coronal mass ejections \citep{crossby93, webb94, lin23}, and an increase in the total solar irradiance \citep{lean1987}. The majority of insights we have gained in long-term activity cycles on other stars is through photometric monitoring and spectroscopic monitoring \citep{saar99}. \cite{lehtinen16} collected 16 to 27 years of B- and V-band photometry for 21 young active solar-type stars. They found evidence of long-term stellar activity cycles in 18 targets following as $P_\textrm{rot}/P_\textrm{cycle} \propto R_0^{-1}$, where $P_\textrm{rot}$ is the rotation period, $P_\textrm{cycle}$ is the period of the stellar activity cycle, and $R_0$ is the Rossby number. Additionally, 50 years of spectroscopic observations of HD~166620 from the Mount Wilson Observatory and Keck revealed both a $\sim 16$~year periodicity in emission from the core of the \ion{Ca}{2} H and K lines \citep{olah16} and evidence that this star entered a grand minima \citep{baum22}. A more complete review on the state of stellar activity cycles can be found by \cite{jeffers23} and \cite{isik23}.

In addition to photometric and spectroscopic observations, solar flares trace the length of the solar activity cycle. The high-cadence observations from \textit{Kepler} (4-year baseline) and TESS (currently 5-year baseline) are providing data sufficient for searching for stellar activity cycle from the variations in stellar flare rates. \cite{davenport20} demonstrated that the flare rate and FFDs of GJ~1234 has not changed appreciably over 10~years of observations with both \textit{Kepler} and TESS. On the other hand, \cite{scoggins19} found that  KIC 8507979, an M3V star, showed a clear decline in flare rate and change in FFD over 4-years of \textit{Kepler} observations. While it is not expected to find flare rate and distribution variations in all stars given detection limitations, KIC 8507979 demonstrates the ability to study long-term flare variability as a potential tracer for stellar activity cycles.

The paper is presented as follows. In Section~\ref{sec:methods}, we describe our sample, stellar flare identification and fitting methods, and methods for measuring stellar rotation periods. In Section~\ref{sec:results}, we present our flare-frequency distribution (FFD) fits as a function of stellar age, $T_\textrm{eff}$, and $R_0$. In Section~\ref{sec:discuss}, we search for correlations in flare rates with FUV and NUV observations from \textit{GALEX} and place the flare rates of young planet hosting stars in the context of our broader sample. We conclude in Section~\ref{sec:conclusions}. We provide additional figures and tables in Section~\ref{appendix:supp_ffds}.

@@

In this section, we provide an overview of the methodology associated with this paper. Specifically, we describe the sample selection in subsection \ref{subsec2:sample}, TESS light curve analysis in \ref{subsec2:TESS_lightcurve}, flare identification in subsection \ref{subsec2:Flareidentification}, how we fit for flare parameters in \ref{subsec2:model}, flare quality checks in \ref{subsec2:qualitychecks}, and measuring stellar rotation periods in \ref{subsec:prot}.

@@

A primary goal of this paper is to measure the dependence of the stellar flare rates on stellar age. Particularly, we are interested in this dependency for stars with ages $1 \leq t_\textrm{age} \leq 300$\,Myr. To this end, we used the MOCA Data Base (Gagné et al. in prep.) to identify nearby young moving groups, associations, and open clusters with known ages which satisfied our criteria. From this database we identified 26~ unique groups from which we created our sample of stars. The final selected targets were required to be: (i) confirmed members, (ii) high-likelihood candidate members, or (iii) candidate members. Membership to these groups has been primarily determined using kinematic information from \textit{Gaia} \citep{gaia16, gaia18}. These rather stringent cuts resulted in a catalog of 30,889 stars across 26 associations. We summarize the sample and ages for each association (and therefore star) in Table~\ref{tab:sample}.

@@

\begin{deluxetable}{l r r r r r}[!ht]
\tabletypesize{\footnotesize}
\tablecaption{Adopted Ages of each Young Stellar Population and Number of
Stars per Group \label{tab:sample}}
\tablehead{
\colhead{Population} & \colhead{Age [Myr]} & \colhead{N$_\textrm{stars}$} &
\colhead{Ref.}
}
\startdata
AB Doradus  &  133$_{-20}^{+15}$  &  88  & 1 \\
Blanco 1 & 137.1$^{+7.0}_{-33}$ & 428 & 2\\
Carina  &  45$\pm 9$  & 94  & 3\\
Carina-Musca & 32 & 35 & 4\\
Chamaeleon  &  5  &  424  & 5 \\
Columba  &  42  &  126  & 3\\
Greater Taurus Subgroup 5 & 8.5 & 56 & 4\\
Greater Taurus Subgroup 8 & 4.5 & 122 & 4\\
Lower Centaurus Crux & 15 & 761 & 6\\
MELANGE-1 & 250$_{-70}^{+50}$ & 19 & 7\\
Octans  &  35$\pm 5$ &   64  & 8 \\
Pisces Eridanis  &  120  &   219  & 9 \\
Pleiades & 127.4$^{+6.3}_{-10}$ & 1421 & 2\\
$\alpha$ Persei  &  79$_{-2.3}^{+1.5}$  &  625 & 2 \\
IC 2602 system &  52.5$_{-3.7}^{+2.2}$  & 160 & 2 \\
NGC 2451A & 48.5 & 59 & 4\\
Oh 59 & 162.2 & 62 & 10\\
Platais 9 & 50 & 124 & 11\\
RSG2 & 126 & 145 & 12\\
Theia 301  &  195  &  437 & 10\\
Theia 95  &  30.2  &  230 & 10\\
TW Hydrae  &  10  &  24  & 3\\
Upper Centaurus Lupus & 16 $\pm$ 2 & 696 & 6\\
Upper Scorpius  &  10  &  106  &  6\\
Vela-CG4  &  33.7  &  299  & 4\\
\hline
Total & & \nstars\\
\enddata
\tablecomments{Age references: (1) \citet{gagne18_abdmg}; (2) \citet{galindo22}; (3) \citet{Bell15}; (4) \citet{kerr21}; (5) \citet{luhman07}; (6) \citet{pecaut16}; (7) \citet{tofflemire21}; (8) \citet{murphy15}; (9) \citet{curtis19}; (10) \citet{kounkel20}; (11) \citet{tarricq21}; (12) \citet{roser16}}
\end{deluxetable}

@@

As described in the previous subsection, we select stars in our sample for which TESS light curves have been obtained with 2-minute cadence. The reason for doing this is so that we are able to resolve and accurately measure the properties of flares for each of these stars \citep{howard22}. We cross-matched our sample obtained with the MOCA Database with the TESS Input Catalog (TIC) based on their Gaia DR2 RA and Dec. We required that the distance between the target and the nearest TIC target was within $<1"$.

This process provided a final sample of \nstars\ unique targets. These targets have each been observed at a 2-minute cadence between TESS Sector 1 and Sector 67; Sector 67 was the latest available sectors at the time that this analysis was performed. In Figure ~\ref{fig:sample}, we show the distribution of the on sky positions and effective temperatures, $T_\textrm{eff}$, of the final sample. We use $T_\textrm{eff}$ as our stellar characterization metric over the standard $M_\star$ because $\sim 600$ stars in our sample are lacking mass information in the TIC.  It is evident that  many targets exhibit strong tendency for proximity towards the
northern and southern ecliptic poles (Figure~\ref{fig:sample}). Therefore, many of these targets were observed
over multiple sectors. The entirety of the available data yielded 17,964 light curves processed by the Science Processing Operations Center (SPOC) pipeline \citep{jenkins16}. Our sample has an average of $\sim 3$ light curves per target (although with significant spread across targets). For our analysis, we used the SPOC-processed SAP\_FLUX.

\begin{figure}[ht!]
    \script{sample.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/sample.pdf}
        \caption{
            Top: Distribution of our selected sample across the sky and colored by the adopted age of the association (see Table~\ref{tab:sample}). The all-sky coverage by TESS has unlocked new populations of stars to observe. We take advantage of this observing strategy to measure flare rates across 26 different nearby young moving groups, clusters, and associations. Bottom: Distribution of adopted effective temperatures, $T_\textrm{eff}$ [K] for stars in our sample. We include all stars with $T_\textrm{eff} \leq 6000$\,K.
        }
        \label{fig:sample}
    \end{centering}
\end{figure}

@@

Once we had the light curves for each target, we performed the following procedure to identify stellar flares. First, we implemented the machine learning flare-identification methods presented and described in \cite{feinstein20}. This method relies on the fact that all flare events have similar time-dependent morphologies. These flare-profiles can generally be described as a sharp rise followed by an exponential decay in the white light curve. This identification-technique implements the convolutional neural network (CNN),
\texttt{stella} \citep{feinstein20}. The neural network was trained to identify flare events in TESS 2-minute data using a catalogue of flare events from TESS Sectors 1 and 2 that were validated  by-eye
\citep{guenther19_flares}.

There are many benefits to using the CNN for stellar flare identification. Primarily, the CNN is advantageous because it is insensitive to the stellar baseline flux. This feature inherently stems from the fact that the network is trained to search \textit{only} based on flare morphology. It is therefore relatively insensitive to the absolute flux levels, so long as the inherent noise does not overwhelm the signal itself. This means that rotational modulation peaks --- which are themselves driven by
stellar heterogeneities --- are not accidentally identified as flares. This is especially advantageous for our sample of young stars, which  readily exhibit rotational modulation in their light curves.

Based on these advantages, the final compiled sample of flares is \textit{unbiased} towards low-amplitude/low-energy flares. It is important to note that these low-energy events are typically not identified in traditional sigma-outlier identification
methods \citep{chang15}. The \texttt{stella} CNN models calculates the probability that a data point in a light curve is associated with a flaring event. Specifically, it takes the light curve (time, flux, flux error) as an input and returns an array with values of [0,1], which are treated as the probability a data point is (1) or is not (0) part of a flare. We ran every light curve through 10 independent \texttt{stella} models and averaged the outputs to ensure that our statistics were accurate. We note that in this processing, the CNN ignores 200-minutes before and after any gaps in the data. Therefore, any flares which occur during these times are not identified.

In order to identify a single flare event, the \texttt{stella} code uses the predictions per data point to group together points. We modified this stage of identification slightly from the original flare-identification method. Specifically, we identified all data points with a probability of being associated with a flare of $P > 0.75$. Any data points that were within 4 cadences of each other were considered to be a single flare event. We did not consider any potential flares that had three or fewer points with $P > 0.75$. This was motivated by the fact that  sometimes a single-point outlier is assigned a high probability of being a flare. Therefore, we neglected these events. We assigned the probability of the whole flare event as the probability of the peak data point.

\begin{figure}[ht!]
    \script{flare_distribution.py}
    \begin{centering}
        \includegraphics[width=0.9\linewidth]{figures/flare_distribution.pdf}
        \caption{
            High level summary of the demographics of flares in our sample. Top: The number of flares identified compared to the number of stars in each nearby young moving group, cluster, or association. A one-to-one relationship is expected. Middle: The distribution measured TESS energies
            and equivalent durations of flares in our sample, colored by the probability of the flare as identified with \texttt{stella}. Bottom: Same as the middle plot, except colored by the $T_\textrm{eff}$ [K] of the star. We limit our sample to stars with $T_\textrm{eff} \leq 6000$\,K.
        }
        \label{fig:flare_distribution}
    \end{centering}
\end{figure}

@@

As described in the previous subsection, flares are well-described in light curves as a sudden increase in flux followed by an exponential decay. For this paper, we used the analytical flare model, Llamaradas Estelares which was presented in
\cite{tovar22}\footnote{\url{https://github.com/lupitatovar/Llamaradas-Estelares}}, to fit the flares and extract the flare parameters in all of the light curves of our sample. This model builds upon the model presented in \cite{davenport14}. Specifically, it includes the convolution of a Gaussian with a double exponential model in the flare profile.

The analytical model is robust in that it is able to accounts for multiple inherent features of flares. Specifically, it can incorporate the flare (i) amplitude, (ii) heating timescale, (iii) rapid cooling phase timescale, and (iv) slow cooling phase timescale. We implemented a nonlinear least squares optimization to fit for the time of the flare peak ($t_\textrm{peak}$), full width at half maximum (FWHM), and the amplitude ($A$) of each flare in the sample.  We combine the model with a second-order polynomial fit to a 1.2-hour baseline before and after the flare  during only the fitting stage. This was implemented in order to account for any slope due to rotational modulation, and therefore was particularly relevant for the rapid rotators ($P_\textrm{rot} < 2$~days).

@@

\begin{equation}\label{eq:Eflare}
    E_\textrm{flare} = L_\star \, A \, \textrm{ED} \, s \, .
\end{equation}
In Equation \ref{eq:Eflare}, $L_\star$ is the luminosity of the star and $s$ is a scaling factor defined as $s = B_\lambda(T_\textrm{eff}) / B_\lambda(T_\textrm{flare})$, where $B_\lambda(T)$ is the Planck function. We assume that the flare temperature is 9000~K \citep{kretzschmar11}. Although 900

@@

\end{document}
feinstein20

>>>>

% Define document class
\documentclass[twocolumn, linenumbers]{aastex631}
\usepackage{showyourwork}
\usepackage{rsfso}
\usepackage{rotating}

@@

% Abstract with filler text
\begin{abstract}

Stellar flares are short-duration ($<$\,hours) bursts of radiation associated with surface magnetic reconnection events. Stellar magnetic activity generally decreases as a
function of both age and Rossby number, $R_0$, a measure of the relative importance of the convective and rotational dynamos.  Young stars ($<300$ Myr) have typically been
overlooked in population-level flare studies because of the lack of long baseline observations and challenges with flare-detection methods. Here, we select a sample of stars
that are members of 26 nearby  moving groups, clusters, or associations with ages $<$300~Myr  that have been observed by the Transiting Exoplanet Survey Satellite (TESS) at
2-minute cadence.  With this sample of \nstars\ stars, we identified \nflares\ flares originating from \nflarestars\ stars. Moreover, we robustly measured the rotation periods
of \nprot~stars within this population. We measure the flare frequency distribution (FFD) slope as a function of both age and spectral type and find the FFD slope, $\alpha$,
saturates for all spectral types at $\alpha \sim -0.5$ and is constant over the 300~Myr. Additionally, we find that flare rates for stars $t_\textrm{age} = 50 - 250$~Myr are
saturated below $R_0 < 0.14$, which is consistent with other indicators of magnetic activity. Taking advantage of the long temporal baseline of TESS, we search for evidence
of annual flare rate variability. We present eleven young stars which tentatively show evidence of flare rate variations across five years of TESS data, potentially correlated
with long term stellar activity cycles. Additionally, we cross match our entire sample  with \textit{GALEX} and find no correlation between flare rate and Far- and Near-Ultraviolet
flux. Finally, we find the flare rates of planet hosting stars are relatively lower than a  comparable, larger samples of stars. This trend may have ramifications for the atmospheric
evolution of young, short-period exoplanets.

@@

Stellar flares are the radiation component of magnetic reconnection events \citep{benz10}. Such events are readily seen on the Sun \citep{carrington1859, lu91, fletcher11},
particularly during the maximum in the solar cycle \citep{webb94}. Solar flares can be used as proxies for magnetic activity occurring on other stars
\citep{feigelson99, berdyugina05, kowalski10, feinstein22_criticality}. Additionally, these short-duration flaring events can have significant ramifications on the evolution
of short-period extrasolar planets \citep{france16, guenther19_flares, chen21}. While stellar flares are typically not spatially resolvable, they do lend themselves to
characterization via spectroscopic and photometric signatures. Spectroscopic characterization of stellar flares inform our understanding of non-thermal processes
affiliated with such events such as coronal mass ejections \citep{argiroffi19, vida19}, proton beams \citep{orrall76, woodgate92}, and accelerated electrons \citep{osten05, smith05}.

Photometric observations of stars are more readily available now with exoplanet transit discovery missions, and allow us to statistically characterize flare rates
and energies at optical/near-infrared wavelengths. Observations of M dwarfs with the Sloan Digital Sky Survey revealed a correlation between the flaring fraction of
stars with height above the galactic plane, a proxy for age \citep{kowalski09, hilton10}. More recently, NASA's \textit{Kepler}, \textit{K2}, and the Transiting
Exoplanet Survey Satellite (TESS) missions have provided a wealth of stellar variability data in addition to their primary objective of detecting transiting exoplanets.
Flares can be identified within time-series photometry by a sharp rise and subsequent exponential decay in flux,  with the latter corresponding to the cooling rate
\citep{kowalski13}. \textit{Kepler} \citep{Borucki10} provided long-baseline high-cadence observations used to identify stellar flares. There has been extensive studies
of flares in \textit{Kepler} data, from the statistics of superflares on solar-type stars \citep[e.g.][]{notsu13, shibayama13, maehara15, okamoto21} to low-mass stars
\citep[e.g.][]{Hawley14, silverberg16}. \cite{davenport19} found that flare activity decreased with increasing rotation period for 347 GKM stars. However, the flare
frequency distribution (FFD) slope did not vary significantly as a function of age. As a caveat, the ages of the stars in this study were determined based on their
rotation periods, relying on the assumption that gyrochronology alone accurately ages stars.

\textit{K2} provided 70-day baseline observations for a handful of young stars in groups such as Upper Scorpius,  Pleiades, Hyades, and Praespe clusters. \cite{ilin19, ilin21}
analyzed flares from K and M stars in these clusters and found that the overall flare activity decreased as a function of age. Moreover, this relationship was steeper for
more massive stars in the sample. \cite{paudel18} surveyed 10 M6 - L0 dwarfs observed with \textit{K2} and found the L0 dwarfs had significantly shallower FFDs than the M
dwarfs. They found that, on average, young targets (defined by the tangential velocity of the star) exhibited more flares overall. More recently, TESS \citep{ricker15} has
provided near all-sky photometric observations at 30-minute cadence or less. This observing strategy has allowed for more detailed studies of young stellar flares from nearby,
disperse young moving groups and associations. These  data permit detailed studies of individual stars, for example characterizing eight superflares ($E_f > 10^{34}$~erg) on
the young star AB Doradus over $\sim 60$ days of continuous observations \citep{schmitt19}, as well as statistical studies of flares across a range of spectral types and ages
\citep{doyle20, feinstein22_criticality, pietras22, yang23}.

The 11-year solar activity cycle represents a change in the magnetic activity of the Sun, and manifests itself in a variety of observables including increases in the total
number of sunspots \citep{clette14, kilcik14}, flares and coronal mass ejections \citep{crossby93, webb94, lin23}, and an increase in the total solar irradiance \citep{lean1987}.
Insights into the long-term activity cycles on other stars have been limited to photometric and spectroscopic monitoring \citep{saar99}. \cite{lehtinen16} collected 16 to 27 years
of B- and V-band photometry for 21 young active solar-type stars. They found evidence of long-term stellar activity cycles in 18 targets for which$P_\textrm{rot}/P_\textrm{cycle}
\propto R_0^{-1}$, where $P_\textrm{rot}$ is the rotation period, $P_\textrm{cycle}$ is the period of the stellar activity cycle, and $R_0$ is the Rossby number. Additionally,
50 years of spectroscopic observations of HD~166620 from the Mount Wilson Observatory and Keck revealed both a $\sim 16$~year periodicity in emission from the core of the
\ion{Ca}{2} H and K lines \citep{olah16} and evidence that this star entered a grand minima \citep{baum22}. A more complete review on the state of stellar activity cycles can be
found in \cite{jeffers23} and \cite{isik23}.

In addition to photometric and spectroscopic observations, solar flares trace the length of the solar activity cycle. The high-cadence observations from \textit{Kepler} (4-year baseline)
and TESS (currently 5-year baseline) provide  sufficient data for searches of stellar activity cycles from the variations in stellar flare rates. \cite{davenport20} demonstrated that
the flare rate and FFDs of GJ~1234 has not changed appreciably over 10~years of observations with both \textit{Kepler} and TESS. On the other hand, \cite{scoggins19} found that the M3V
star star KIC 8507979 showed a clear decline in flare rate and change in FFD over 4-years of \textit{Kepler} observations. While it is not expected to find flare rate and distribution
variations in all stars given detection limitations, KIC 8507979 demonstrates the ability to study long-term flare variability as a potential tracer for stellar activity cycles.

The paper is presented as follows. In Section~\ref{sec:methods}, we describe our sample, and stellar flare and rotation period identification methods. In Section~\ref{sec:results},
we present our flare-frequency distribution (FFD) fits as a function of stellar age, $T_\textrm{eff}$, and $R_0$. In Section~\ref{sec:cycles}, we present evidence of flare rate changes
over the 5-year TESS baseline in eleven young stars, likely correlated with long term stellar activity cycles. In Section~\ref{sec:discuss}, we search for correlations in flare rates
with FUV and NUV observations from \textit{GALEX} and place the flare rates of young planet hosting stars in the context of our broader sample.  We conclude in Section~\ref{sec:conclusions}.
We provide additional figures and tables in Section~\ref{appendix:supp_ffds}.

@@

Here, we provide an overview of the methodology used in this paper. Specifically, we describe the sample selection in Section~\ref{subsec2:sample}, TESS light curve analysis in
Section~\ref{subsec2:TESS_lightcurve}, flare identification in Section~\ref{subsec2:Flareidentification}, flare fitting parameters in Section~\ref{subsec2:model}, flare quality
checks in Section~\ref{subsec2:qualitychecks}, and  stellar rotation period measurements in Section~\ref{subsec:prot}.

@@

A primary goal of this paper is to measure the relationship between the flare rates and ages of stars. We are particularly interested in this dependency for young stars with ages
$4 \leq t_\textrm{age} \leq 300$\,Myr. To this end, we used the MOCA Data Base (Gagné et al. in prep.)\footnote{\url{https://mocadb.ca/}} to identify 26~ nearby, aged, young moving
groups, associations, and open clusters from which we created our sample of stars. The final targets were required to be: (i) confirmed members, (ii) high-likelihood candidate members,
or (iii) candidate members.  Membership to these groups has been primarily determined using kinematic information from \textit{Gaia} \citep{gaia16, gaia18}. The membership status is
determined by the probability that BANYAN-$\Sigma$ assigns based on how well the kinematics of the target matches with the kinematics of the group \citep{gagne18}. These rather
stringent cuts resulted in a catalog of 30,889 stars across 26 associations. We summarize the sample and ages for each association (and therefore star) in Table~\ref{tab:sample}.

@@

\begin{deluxetable}{l r r r r r}[!ht]
\tabletypesize{\footnotesize}
\tablecaption{Adopted Ages of each Young Stellar Population and Number of
Stars per Group, $N_\textrm{stars}$ \label{tab:sample}}
\tablehead{
\colhead{Population} & \colhead{Age [Myr]} & \colhead{N$_\textrm{stars}$} &
\colhead{Ref.}
}
\startdata
AB Doradus  &  133$_{-20}^{+15}$  &  88  & 1 \\
Blanco 1 & 137.1$^{+7.0}_{-33}$ & 428 & 2\\
Carina  &  45$\pm 9$  & 94  & 3\\
Carina-Musca & 32 & 35 & 4\\
Chamaeleon  &  5  &  424  & 5 \\
Columba  &  42  &  126  & 3\\
Greater Taurus Subgroup 5 & 8.5 & 56 & 4\\
Greater Taurus Subgroup 8 & 4.5 & 122 & 4\\
Lower Centaurus Crux & 15 & 761 & 6\\
MELANGE-1 & 250$_{-70}^{+50}$ & 19 & 7\\
Octans  &  35$\pm 5$ &   64  & 8 \\
Pisces Eridanis  &  120  &   219  & 9 \\
Pleiades & 127.4$^{+6.3}_{-10}$ & 1421 & 2\\
$\alpha$ Persei  &  79$_{-2.3}^{+1.5}$  &  625 & 2 \\
IC 2602 system &  52.5$_{-3.7}^{+2.2}$  & 160 & 2 \\
NGC 2451A & 48.5 & 59 & 4\\
Oh 59 & 162.2 & 62 & 10\\
Platais 9 & 50 & 124 & 11\\
RSG2 & 126 & 145 & 12\\
Theia 301  &  195  &  437 & 10\\
Theia 95  &  30.2  &  230 & 10\\
TW Hydrae  &  10  &  24  & 3\\
Upper Centaurus Lupus & 16 $\pm$ 2 & 696 & 6\\
Upper Scorpius  &  10  &  106  &  6\\
Vela-CG4  &  33.7  &  299  & 4\\
\hline
Total & & \nstars\\
\enddata
\tablecomments{Age references: (1) \citet{gagne18_abdmg}; (2) \citet{galindo22}; (3) \citet{Bell15}; (4) \citet{kerr21}; (5) \citet{luhman07}; (6) \citet{pecaut16}; (7)
\citet{tofflemire21}; (8) \citet{murphy15}; (9) \citet{curtis19}; (10) \citet{kounkel20}; (11) \citet{tarricq21}; (12) \citet{roser16}}
\end{deluxetable}

@@

We cross-matched our MOCA Database sample with the TESS Input Catalog (TIC) based on their Gaia DR2 RA and Dec; we required that the distance between the target and the nearest
TIC target was within $<1"$. We down-selected our sample to stars that have been observed with TESS at 2-minute cadence. This ensures that we are able to temporally resolve and
accurately measure the properties of flares for each of these stars \citep{howard22}.

This process provided a final sample of \nstars\ unique targets. These targets have each been observed at a 2-minute cadence between TESS Sector 1 and Sector 67; Sector 67 was
the latest available sector at the time that this analysis was performed. In Figure ~\ref{fig:sample}, we show the distribution of the on sky positions, ages and effective
temperatures, $T_\textrm{eff}$, of the final sample. Because many of the groups are located at or near the ecliptic poles, many of our targets were observed over multiple TESS
sectors. We use $T_\textrm{eff}$ as our stellar characterization metric over the standard $M_\star$ because $\sim 600$ stars in our sample lack mass information in the TIC.
The entirety of the available data yielded 17,964 light curves processed by the Science Processing Operations Center (SPOC) pipeline \citep{jenkins16}. Our sample has an average
of $\sim 3$ light curves per target (although with significant spread across targets). For our analysis, we used the SPOC-processed SAP\_FLUX.

\begin{figure}[ht!]
    \script{sample.py}
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/sample.pdf}
        \caption{
            Top: Distribution of the selected sample across the sky and colored by the adopted age of the association (see Table~\ref{tab:sample}). The all-sky coverage by TESS
            has unlocked new populations of stars to observe. We take advantage of this observing strategy to measure flare rates across 26 different nearby young moving groups,
            clusters, and associations. Bottom: Distribution of adopted effective temperatures, $T_\textrm{eff}$ [K] for stars in our sample. We include all stars with $T_\textrm{eff} \leq 6000$\,K.
        }
        \label{fig:sample}
    \end{centering}
\end{figure}

@@

Once we downloaded the light curves for each target, we performed the following procedure to identify stellar flares. First, we implemented the machine learning flare-identification
methods presented and described in \cite{feinstein20}. This method relies on the similar time-dependent morphologies of all flare events. These flare-profiles can generally be
described as a sharp rise followed by an exponential decay in the white light curve. This identification-technique implements the convolutional neural network (CNN) \texttt{stella}
\citep{feinstein20}. The CNN  was trained on  a   by-eye validated catalogue of flares from TESS Sectors 1 and 2 with 2-minute data \citep{guenther19_flares}.

There are several benefits to using the CNN for stellar flare identification. Primarily, the CNN is insensitive to the stellar baseline flux because it is trained to search
\textit{only} based on flare morphology. It is therefore relatively insensitive to the absolute flux levels, so long as the inherent noise does not overwhelm the signal itself.
An additional benefit is that rotational modulation peaks --- which are themselves driven by stellar heterogeneities --- are not accidentally identified as flares. This holds true
for stars with rotation periods, $P_\textrm{rot} > 1$~day. This is especially advantageous for our sample of young stars, which readily exhibit rotational modulation in their light curves.

Based on these advantages, the final compiled sample of flares is unbiased towards low-amplitude/low-energy flares. It is important to note that these low-energy events are typically
not identified in traditional sigma-outlier identification
methods \citep{chang15}. The \texttt{stella} CNN models calculate the probability that a data point in a light curve is associated with a flaring event. Specifically, it takes the light
curve (time, flux, flux error) as an input and returns an array with values of [0,1], which are treated as the probability a data point is (1) or is not (0) part of a flare. We ran every
light curve through 10 independent \texttt{stella} models and averaged the outputs to ensure that our statistics were accurate. We note that in this processing, the CNN ignores 200-minutes
before and after any gaps in the data. Therefore, any flares which occur during these times are not identified.

The \texttt{stella} code groups individual points with the predictions per data point when identifying a single flare event. We modified this stage of identification slightly from the
original flare-identification method. Specifically, we identified all data points with a probability of being associated with a flare of $P > 0.75$. Any data points that were within 4
cadences of each other were considered to be a single flare event. We did not consider any potential flares that had three or fewer points with $P > 0.75$. This method rejects single-point
outliers which can be assigned  high probabilities of being  flares. We assigned the probability of the whole flare event as the probability of the peak data point.

\begin{figure}[ht!]
    \script{flare_distribution.py}
    \begin{centering}
        \includegraphics[width=0.9\linewidth]{figures/flare_distribution.pdf}
        \caption{
            High level summary of the demographics of flares in our sample. Top: The number of flares identified compared to the number of stars in each nearby young moving group, cluster,
            or association. A linear relationship is expected. Bottom: The distribution of measured TESS energies
            and equivalent durations of flares in our sample, colored by the probability of the flare as identified with \texttt{stella}.
        }
        \label{fig:flare_distribution}
    \end{centering}
\end{figure}

@@

Flares are well-described in light curves as a sudden increase in flux followed by an exponential decay. We used the analytic flare model, Llamaradas Estelares which was presented in
\cite{tovar22}\footnote{\url{https://github.com/lupitatovar/Llamaradas-Estelares}}, to fit and extract parameters of flares in all of the light curves of our sample. This model builds
upon the model presented in \cite{davenport14}. Specifically, it includes the convolution of a Gaussian with a double exponential model in the flare profile.

The analytic model robustly  accounts for physically-motivated flare features. Specifically, it can incorporate the flare (i) amplitude, (ii) heating timescale, (iii) rapid cooling phase
timescale, and (iv) slow cooling phase timescale. We implemented a nonlinear least squares optimization to fit the flare peak time ($t_\textrm{peak}$), full width at half maximum (FWHM),
and the amplitude ($A$) of each flare in the sample.  We combine the model with a second-order polynomial fit to a 1.2-hour baseline before and after the flare  during only the fitting stage.
This was implemented in order to account for any slope due to rotational modulation, and therefore was particularly relevant for the rapid rotators ($P_\textrm{rot} < 2$~days).

@@

\begin{equation}\label{eq:Eflare}
    E_\textrm{flare} = L_\star \, A \, \textrm{ED} \, s \, .
\end{equation}
In Equation \ref{eq:Eflare}, $L_\star$ is the luminosity of the star and $s$ is a scaling factor defined as $s = B_\lambda(T_\textrm{eff}) / B_\lambda(T_\textrm{flare})$, where $B_\lambda(T)$ is
the Planck function. We assume that the flare temperature is 9000~K \citep{hawley92, hawley95}, although recent NUV flare observations suggest this may be an underestimation
\citep{kowalski19, brasseur23, berger23}.

@@

\end{document}
